MSCA 31006: Final Project - AQI Forecasting

### 2023-08-13 MSCA_31006_Time_Series_Final_Project - TBATS.Rmd
Add files via upload

```{r, tidy=TRUE}
path <- "/Users/aashishsingh/Desktop/Time Series - Final Project/"
pm2_5_data <- read.csv(paste0(path, "la_pm2_5_pollutant_data.csv"), 
                     na.strings=c("", "NA"))
ozone_data <- read.csv(paste0(path, "la_ozone_pollutant_data.csv"), 
                     na.strings=c("", "NA"))

@@

>>>>

```{r, tidy=TRUE}
#path <- "/Users/aashishsingh/Desktop/Time Series - Final Project/"
#pm2_5_data <- read.csv(paste0(path, "la_pm2_5_pollutant_data.csv"), 
#                     na.strings=c("", "NA"))
#ozone_data <- read.csv(paste0(path, "la_ozone_pollutant_data.csv"), 
#                     na.strings=c("", "NA"))

path <- "D:/TS Final/"
pm2_5_data <- read.csv(paste0(path, "pm2.5_LA.csv"), 
                     na.strings=c("", "NA"))
ozone_data <- read.csv(paste0(path, "ozone_LA.csv"), 
                     na.strings=c("", "NA"))

@@

#### Step 9a: TBATS model - Weekly

```{r, tidy=TRUE}
vec <- as.vector(train_weekly)

univariate_ts <- ts(vec, start=1999, frequency=52.18)

stl_decomposition <- stl(univariate_ts, s.window="periodic")
plot(stl_decomposition)

h <- length(test_weekly)

model_tbats <- tbats(train_weekly)
forecast_tbats <- forecast(model_tbats, h=h)
plot(forecast_tbats)

checkresiduals(model_tbats)
#Ljung-Box test .05179

plot(pm2_5_weekly_ts, xlim=c(1999, 2023), ylim=c(min(pm2_5_weekly_ts), max(pm2_5_weekly_ts)), main="Train, Test, and Forecasted Data - Weekly", xlab="Year", ylab="PM2.5 AQI Value")
lines(test_weekly, col="red") # Test data in red
lines(forecast_tbats$mean, col="blue", type="o") # Forecasted data in blue
legend("topleft", legend=c("Train", "Test", "Forecast"), fill=c("black", "red", "blue"))

mae <- mean(abs(test_weekly - forecast_tbats$mean))
print(cat("Mean Absolute Error (MAE):", round(mae, 2), "\n"))
#Mean Absolute Error (MAE): 9.31

comparison_df <- data.frame(
  Date = time(test_weekly),
  Actual = as.vector(test_weekly),
  Forecasted = as.vector(forecast_tbats$mean)
)
print(comparison_df)

#### Step 9b: TBATS model - Monthly

vec <- as.vector(train_monthly)

univariate_ts <- ts(vec, start=1999, frequency=52.18)

stl_decomposition <- stl(univariate_ts, s.window="periodic")
plot(stl_decomposition)

h <- length(test_monthly)

model_tbats <- tbats(train_monthly)
forecast_tbats <- forecast(model_tbats, h=h)
plot(forecast_tbats)

checkresiduals(model_tbats)
#Ljung-Box test .07956

plot(pm2_5_monthly_ts, xlim=c(1999, 2023), ylim=c(min(pm2_5_monthly_ts), max(pm2_5_monthly_ts)), main="Train, Test, and Forecasted Data - Monthly", xlab="Year", ylab="PM2.5 AQI Value")
lines(test_monthly, col="red") # Test data in red
lines(forecast_tbats$mean, col="blue", type="o") # Forecasted data in blue
legend("topleft", legend=c("Train", "Test", "Forecast"), fill=c("black", "red", "blue"))

mae <- mean(abs(test_monthly - forecast_tbats$mean))
print(cat("Mean Absolute Error (MAE):", round(mae, 2), "\n"))
#Mean Absolute Error (MAE): 6.61

comparison_df <- data.frame(
  Date = time(test_monthly),
  Actual = as.vector(test_monthly),
  Forecasted = as.vector(forecast_tbats$mean)
)
print(comparison_df)
